---
title: "Budweiser Analysis"
author: "Edward Roske"
date: "12/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Presentation for Budweiser leadership
* **CEO:** Carlos Brito
* **CFO:** Felipe Dutra  

# Craft Beer Analysis
## This presentation is to report and analyze craft beer data to identify potential options for Budweiser's strategic direction.  

***


```{r header, echo=FALSE, collapse=TRUE, include=FALSE}
# MSDS6306 - Unit 8
# Edward Roske
# 12/15/2019

# Tidyverse
library (tidyverse)

# Graphics
library (ggplot2)
library (ggthemes)
library (usmap)

# XML, JSON, strings
library (rvest) #html_table, html_node
library (stringi)
library (XML) #xml_Parse
library (httr)
library (plyr)
library (tm) #text mining library provides the stopwords() function
library (tidyr)
library (jsonlite)
library (dplyr)

# Statistics
library (class)
library (caret)
library (e1071) #Naive Bayes
```

## Begin by loading the data and preparing it

```{r prepare}
# Unit 8: Beer Analysis for Budweiser

# Read dataset
# Some fields (State in Breweries) start with spaces, so remove them
url <- "https://raw.githubusercontent.com/ERoske/MSDS-6306/master/"
beers <- read.csv(paste0(url,"Beers.csv"), strip.white=TRUE, sep=",")
breweries <- read.csv(paste0(url,"Breweries.csv"), strip.white=TRUE, sep=",")
beers$Ounces <- as.factor(beers$Ounces)

# Add variables
# Ale, IPA, or neither
beers$Ale <- "No"
beers$Ale[grep("Ale", beers$Style)] <- "Other"
beers$Ale[grep("IPA", beers$Style)] <- "IPA"
```

***

## **Question 3:**
## Address the missing values in each column.
## Missing values: ABV, IBU, Style
## Should we leave empty or fill with averages?

```{r Q3}
sum(is.na(beers$ABV))
sum(is.na(beers$IBU))
sum(beers$Style=="")
```

## **Answer 3:**
## For now, we will replace with overall medians (or generic "Other").  

```{r A3}
beers$ABV[is.na(beers$ABV)] <- median(beers$ABV, na.rm=TRUE)
beers$IBU[is.na(beers$IBU)] <- median(beers$IBU, na.rm=TRUE)
beers$Style[beers$Style==""] <- "Other"
```
***
## **Question 1:**
## How many breweries are present in each state?

```{r Q1}
state_table <- as.data.frame(table(breweries$State))
state_summary <- merge(statepop, state_table, by.x="abbr", by.y="Var1")
state_summary$PerCapita <- round(state_summary$pop_2015 / state_summary$Freq)

```

## **Answer 1:**
## Here is the data summarized by State (and District of Columbia):

```{r A1}
# Table: Brewery Count
knitr::kable(state_table,
    caption = "Breweries present in each state")

# Graph: Brewery Count
plot_usmap(data=state_summary, values="Freq", color="black") +
    scale_fill_continuous(
        low="white", high="red", name="Breweries", label=scales::comma
    ) +
    theme(legend.position="right") +
    labs(title="Brewery Analysis by State") +
    theme(plot.title=element_text(size=18, hjust=0.5, face="bold"))
```

***

## **Question 2:**
## Merge beer data with the breweries data.  

``` {r Q2}
merged <- merge(beers, breweries, by.x="Brewery_id", by.y="Brew_ID")
names(merged)[1] <- 'Brewery_ID'
names(merged)[2] <- 'Beer_Name'
names(merged)[9] <- 'Brewery_Name'

# Print first 6 and last 6 rows of the merged file
#head(merged, n=6)
#tail(merged, n=6)

data <- rbind(head(merged, n=6),
          tail(merged, n=6))
```

## **Answer 2:**
## Here are the first 6 and last 6 rows of the merged file.

``` {r A2}
knitr::kable(data,
    caption = "Merged Beer/Brewery Data")
```

***

## **Question 4:**
## Compute the median alcohol content and IBU for each state.
## Print and graph the ABV and IBU for each state.

``` {r Q4}
state_medians <-
    merged %>%
    group_by(State) %>%
    dplyr::summarise(IBU = median(IBU),
                     ABV = median(ABV))
state_summary <- merge(state_summary, state_medians, by.x="abbr", by.y="State")

## Print a table with ABV and then IBU for each state.
knitr::kable(state_summary,
    caption = "Median Alcohol Content and IBU for each state")

# Graph IBUs and ABVs
# IBU
ggplot(state_summary,
       aes(x=abbr,
           y=IBU,
           fill=IBU)) +
    geom_bar(stat="identity") +
    scale_fill_gradient(low="blue", high="red") +
    labs(x="States",
         y="IBU",
         title="IBU Analysis by State") +
    theme_economist_white() +
    theme(legend.position="none",
          plot.title=element_text(size=18, hjust=0.5, face="bold"),
          axis.title.x=element_text(size=12, face="bold"),
          axis.title.y=element_text(size=12, face="bold"),
          axis.text.x=element_text(size=10),
          axis.text.y=element_text(size=6),
    ) +
    coord_flip()
# ABV
ggplot(state_summary,
       aes(x=abbr,
           y=ABV,
           fill=ABV)) +
    geom_bar(stat="identity") +
    scale_fill_gradient(low="blue", high="red") +
    labs(x="States",
         y="IBU",
         title="ABV Analysis by State") +
    theme_economist_white() +
    theme(legend.position="none",
          plot.title=element_text(size=18, hjust=0.5, face="bold"),
          axis.title.x=element_text(size=12, face="bold"),
          axis.title.y=element_text(size=12, face="bold"),
          axis.text.x=element_text(size=10),
          axis.text.y=element_text(size=6),
    ) +
    coord_flip()
```

## **Answer 4:**
## For IBU, the highest states are West Virginia, Mississippi, Delaware, and Alabama.
## For ABV, there is much more variance, but the highest states are District of Columbia, West Virginia, Alabama, and Kentucky.
## In general, the Southern states tend to have higher bitterness and alcohol content.

***

## **Question 5:**
## Which state has the maximum alcoholic (ABV) beer?
## Which state has the most bitter (IBU) beer?

``` {r A5}
paste("Top ABV State:",
    state_summary[order(-state_summary$ABV),"abbr"][1])
paste("Top IBU State:",
      state_summary[order(-state_summary$IBU),"abbr"][1])
```

## **Answer 5:**
## Assuming they want maximum medians from Question 4 and not a single beer outlier,
* The top ABV state is DC.
* The top IBU state is WV.  

***

## **Question 6:**
## Comment on the summary statistics and distribution of the ABV variable.

``` {r Q6}
summary(merged$ABV)
sd(merged$ABV)
ggplot(merged,
       aes(x=Ounces,
           y=ABV,
           fill=Ounces)) +
    geom_boxplot() +
    labs(x="Ounces",
         y="ABV",
         title="ABV Analysis by Ounces") +
    scale_y_continuous(name="ABV", breaks=seq(0.0,0.15,by=.01)) +
    theme_economist_white() +
    theme(legend.position="none",
          plot.title=element_text(size=18, hjust=0.5, face="bold"),
          axis.title.x=element_text(size=12, face="bold"),
          axis.title.y=element_text(size=12, face="bold"),
          axis.text.x=element_text(size=10),
          axis.text.y=element_text(size=10),
    )
```

## **Answer 6:**
## The median ABV is 5.60%. The mean is 5.97% implying a right-skewed distribution.
## The interquartile range is fairly tight ranging from 5.0%-6.7%.
## 12-Ounce beers tend to have the lowest average ABV.
## 19.2-Ounce beers tend to have the highest average ABV.

***

## **Question 7:**
## Is there an apparent relationship between the bitterness of the beer and its alcoholic content?
## Draw a scatter plot.
## Make your best judgment of a relationship and EXPLAIN your answer.

``` {r Q7}
# All Styles
MergedClean <- merged %>%
    filter(!ABV==median(beers$ABV, na.rm=TRUE) &
           !IBU==median(beers$IBU, na.rm=TRUE))
ggplot(MergedClean, aes(x=IBU, y=ABV, color=Ale)) +
    scale_x_continuous(name="IBU", breaks=seq(0,140,by=10), limits=c(0, 140)) +
    scale_y_continuous(name="ABV", breaks=seq(0,0.13,by=.01), limits=c(0, 0.13)) +
    ggtitle("ABV vs IBU Analysis") +
    geom_point(position="jitter") +
    geom_smooth(color="red", size=2, se=FALSE, method="lm") +
    #scale_y_continuous(name="ABV", breaks=seq(0.0,0.15,by=.01)) +
    theme_economist_white() +
    theme(plot.title=element_text(size=18, hjust=0.5, face="bold"),
          axis.title.x=element_text(size=12, face="bold"),
          axis.title.y=element_text(size=12, face="bold"),
          axis.text.x=element_text(size=10),
          axis.text.y=element_text(size=10),
    )
# Ale comparisons
Ales <- MergedClean %>% filter(!Ale=="No")
ggplot(data=Ales, aes(x=IBU, y=ABV, color=Ale)) +
    scale_x_continuous(name="IBU", breaks=seq(0,140,by=10), limits=c(0, 140)) +
    scale_y_continuous(name="ABV", breaks=seq(0,0.13,by=.01), limits=c(0, 0.13)) +
    ggtitle("ABV vs IBU Analysis") +
    geom_point(position="jitter") +
    geom_smooth(size=2, se=FALSE, method="lm") +
    #scale_y_continuous(name="ABV", breaks=seq(0.0,0.15,by=.01)) +
    theme_economist_white() +
    theme(plot.title=element_text(size=18, hjust=0.5, face="bold"),
          axis.title.x=element_text(size=12, face="bold"),
          axis.title.y=element_text(size=12, face="bold"),
          axis.text.x=element_text(size=10),
          axis.text.y=element_text(size=10),
    )
```

## **Answer 7:**
## There is a direct relationship (positive correlation according to the graph) of IBU and ABV.
## In general, the higher the IBU (Bitterness Units), the higher the ABV.
## For IPAs, it is even more impactful that as IBUs increase, ABV increases even more.

***

## **Question 8:**
## Investigate the difference with respect to IBU and ABV between Ales (IPA and Others)
## use KNN classification to investigate this relationship.

``` {r Q8}
# kNN Classification
knnClassifications <-
    knn.cv(Ales[,4:5],
           Ales$Ale,
           k=round(sqrt(nrow(Ales))))
(knnCM <- confusionMatrix(table(knnClassifications, Ales$Ale)))
```

## **Answer 8:**
## Based on K-NN Analysis, we can predict with greater than 81% Accuracy
## if something is an IPA or another Ale based on IBU & ABV.

****

## **Question 9:**
## Find one other useful inference.
## Graph: Per Capita

``` {r Q9}
plot_usmap(data=state_summary, values="PerCapita", color="black") +
    scale_fill_continuous(
        low="red", high="green", name="Population/Breweries", label=scales::comma
    ) +
    theme(legend.position="right") +
    labs(title="Brewery Analysis by State:",
         subtitle="Population per Brewery as of 2015") +
    theme(plot.title=element_text(size=18, hjust=0.5, face="bold"),
          plot.subtitle=element_text(size=18, hjust=0.5),
    )
```

## **Answer 9:**
## Red areas are oversaturated (too many breweries per capita).
## Green areas need new craft breweries.
## I advise that we buy existing craft breweries (or open new ones) in 
## Tennessee and New Jersey.
## Other possibilities could include West Virginia and Nevada.

***

## **Observation**
## This data seems to be from a few years ago (2015?). In the years since, 
## New Jersey has seen a major increase in craft breweries. Here's an Interesting article:
<https://www.northjersey.com/story/life/food/2019/01/03/north-jerseys-craft-brewery-scene-has-grown-more-than-any-other-state/2474070002/>

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
